Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-01-29 paper/introduction.tex
pass over introduction

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common tasks are captured as parameterized templates.
%
%More powerful still are systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
%and model finding tools to fill in parameters in configurations while
%ensuring key correctness constraints are satisfied.
One step further, systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure} use SAT solving
and model finding tools to correctly and consistently fill in some parameters.
%
While templates help ensure certain kinds of consistency across devices,
they do not bridge the
semantic divide between network-wide policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages and so still require operators to
manually decompose policies into per-device behaviors.

@@

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, letting users specify paths within their network,
they fail to provide a coherent means to specify preferred
\emph{inter}-domain routes.  \todd{What is the impact of these
  languages not supporting inter-domain routing?  How are those
  policies expressed instead in an SDN context?  A sentence on the
  impact would help.}  Second, centralized control planes
require careful design and engineering to be robust to failures---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of large, geographically distributed networks,
and to react quickly
to a changing environment. To address this challenge, researchers are exploring
multi-controller systems with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

@@

To achieve the first goal, we borrow the idea of using regular
expressions to specify network paths from 
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, our design also contains several key
departures from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the exact same way.
%No SDN programming language provides similar facilities.  
In addition, we augment the algebra
of regular expressions to directly support a notion of {\em preferences} and provide a semantics in terms of sets of ranked paths.  The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the highest-ranked one will be taken.  
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

To achieve the second goal, we develop program analysis and compilation
algorithms that translate the regular policies to a graph-based
intermediate language and from there to BGP configurations, including
per-device import and export filters, local preferences, MED attributes, and community tags.
We target BGP for pragmatic reasons:  it is an industry
standard and many networks use it internally as well as externally. Despite the advent of SDN, many networks will continue to
use BGP for the foreseeable future due to existing infrastructure investments, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.

The BGP configurations produced by our compiler are
guaranteed to satisfy the given policy specification in the face of
{\em arbitrary} failures.  In this way, we provide network operators
with a strong guarantee that is otherwise impossible to achieve
today. 
However, some policies simply cannot be implemented correctly in BGP in 
 the presence of arbitrary failures.  We develop new
algorithms to detect such policies
and report our findings to the operators, so they may fix the policy
specification at compile time rather than experience undesirable
behavior after the configurations are deployed.
%We also analyze route aggregation to detect the possibility of failure-induced black-holes.

>>>>

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common tasks are captured as parameterized templates.
%
%More powerful still are systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
%and model finding tools to fill in parameters in configurations while
%ensuring key correctness constraints are satisfied.
%One step further, systems like
%ConfigAssure~\cite{narain:lisa05,narain+:configassure} use SAT solving
%and model finding tools to correctly and consistently fill in some parameters.
%
While templates help ensure certain kinds of consistency across devices,
they do not bridge the
semantic divide between network-wide policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages and so still require operators to
manually decompose policies into device-level behaviors.

@@

The centralized control planes of SDN, however, are not a panacea.
First, while many SDN systems provide effective \emph{intra}-domain routing
abstractions, letting users specify paths within their network,
they fail to provide a coherent means to specify preferred
\emph{inter}-domain routes. 
This means that existing inter-domain routing protocols and their associated configuration difficulties will be around for the forseeable future.
\ryan{I added this. Is it true? what about SDX?}
\todd{What is the impact of these
  languages not supporting inter-domain routing?  How are those
  policies expressed instead in an SDN context?  A sentence on the
  impact would help.}  
Second, centralized control planes
require careful design and engineering to be robust to failures---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of large, geographically distributed networks,
and to react quickly
to a changing environment. To address this challenge, researchers are exploring
multi-controller systems with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.

@@

To achieve the first goal, we borrow the idea of using regular
expressions to specify network paths from 
recent high-level SDN languages such as FatTire~\cite{fattire},
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, our design also contains several key
departures from existing languages.  The most important difference is semantic:  the paths specified
by programmers can extend from outside the operator's network to inside
the network, across several devices internally, and then out again.  This design
choice allows users to specify preferences about both external and internal
routes in the exact same way.
%No SDN programming language provides similar facilities.  
In addition, we augment the algebra
of regular expressions to directly support a notion of {\em preferences} and provide a semantics in terms of sets of ranked paths. The preferences indicate fail-over behaviors:  among all specified paths that are still available,
the system guarantees that the distributed implementation will always use the highest-ranked ones.  
%We also support network operators by defining a
%set of common abbreviations for entering, leaving, and traversing
%the user network and we provide abstractions for managing common features %of
%control plane algorithms such as route aggregation.  Operators use familiar logical operations
%to build routes in a modular and compositional manner from these
%primitives.  Our abbreviations are translated in to our regular-expression-based
%intermediate representation and compiled from there.

To achieve the second goal, we develop program analysis and compilation
algorithms that translate the regular policies to a graph-based
intermediate language and from there to BGP configurations, including
per-device import and export filters, local preferences, MED attributes, and community tags.
We target BGP for pragmatic reasons: it is an industry
standard, and many networks use it internally as well as externally. Despite the advent of SDN, many networks will continue to
use BGP for the foreseeable future due to existing infrastructure investments, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.

The BGP configurations produced by our compiler are
guaranteed to be policy compliant in the face of
{\em arbitrary} failures.  In this way, we provide network operators
with a strong guarantee that is otherwise impossible to achieve
today. 
However, some policies simply cannot be implemented correctly in BGP in 
 the presence of arbitrary failures.  We develop new
algorithms to detect such policies
and report our findings to the operators, so they may fix the policy
specification at compile time rather than experience undesirable
behavior after the configurations are deployed.
%We also analyze route aggregation to detect the possibility of failure-induced black-holes.
